---
title: "心統實習課 - 描述統計(繪圖)"
author: "WC_Chang"
date: "2023-10-02"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

# 上週今日

-   路徑(path)、工作位置(working directory)和匯入/匯出檔案
    -   `getwd()`
    -   `setwd()`
-   描述統計的一堆函數
-   有點挑戰的作業

## 上週作業和匯出檔案的補充

-   上週作業其實有很多種作法

```{r }
dta1 <- read_csv("C:/Users/wcnon/OneDrive/Desktop/Graduate/Statistics_TA/RLab/03_Descriptive_Statistics/fake_data.csv")

## 上週作業
avg <- c(mean(dta1[dta1$gender == "M", ]$quiz1, na.rm = T),
          mean(dta1[dta1$gender == "F", ]$quiz1, na.rm = T)
         )
sd <- c(sd(dta1[dta1$gender == "M", ]$quiz1, na.rm = T),
        sd(dta1[dta1$gender == "F", ]$quiz1, na.rm = T))
des <- data.frame(gender = c("M","F"),
                  avg.quiz1 = avg,
                  sd.quiz1 = sd)
des
```

### 匯出資料

-   兩種方法

    -   `base` 語法:`write.table()`, `write.csv()`

    -   `readr`套件: `write_csv()`

```{r output, warning=FALSE}
setwd("C:/Users/wcnon/OneDrive/Desktop/")

#大家好像都抄這個:(
write.table(des, file="output.csv",
            col.names = T, row.names = T, sep = ",")

write.table(des, file="output1.csv",
            col.names = T, row.names = F, sep = ",")
#另一個
write.csv(des, file="output2.csv")
#第三個
readr::write_csv(des, file="output3.csv")

#txt檔如果你喜歡的話
write.table(des, file="output.txt",
            col.names = T,
            sep = ",")
```

# 本周要用的套件

`ggplot2`

```{r }
#install.packages("tidyverse")
#install.packages("ggplot2")
library(ggplot2) 
```

# Iris 資料集

先看看 `iris` 長怎樣
```{r take a peep of iris}
?iris
str(iris)
head(iris)
colnames(iris)
```

# My first plot in ggplot2

-   3 Major components:
    1.  data, should be `data.frame` format.
    2.  A set of **aesthetic mappings** between variables in the data and visual properties, and
    3.  At least one layer which describes how to render each observation. Layers are usually created with a `geom` function.


For Example: 

```{r little example}
ggplot(data = iris, mapping = aes(x = Sepal.Length, y = Sepal.Width)) +
  geom_point()  #layer_2: type of plot 
```

## On `aes()`: Colour, shape and size

Try these:

-   `aes(Sepal.Length, Sepal.Width, color = Species)`

-   `aes(Sepal.Length, Sepal.Width, shape = Species)`

# Outline---Single variable X

| 類別變數 (Categorical var.) | 連續變數 (Continuous var.) |
|:----------------------------|:---------------------------|
|                             | 直方圖(histogram)          |
| 長條圖(bar plot)            | 盒鬚圖(box plot)           |
| ~~圓餅圖(pie chart)~~       | ~~QQ圖(QQ plot)~~          |

## Outline---Multivariate (x,y)

| 資料類型     | 圖                                 |
|--------------|------------------------------------|
| 二個類別     | 通常用關聯表 (contingency table)   |
| 一類別一連續 | 平行盒鬚圖 (side‐by‐side box plot) |
| 二個連續     | 散佈圖 (scatter plot)              |
| 三個以上變項 | 以上圖形搭配不同顏色、形狀         |

## bar chart

```{r iris bar}
ggplot(iris[iris$Sepal.Length>5,], aes(Species))+
   geom_bar(fill = "darkcyan", width = .5) +
   labs(title="Observations of Sepal.Length > 5")
```

## histogram

```{r 2}
ggplot(iris, aes(Sepal.Width)) +
  geom_histogram()+
  labs(title = "default(bins = 30)")# bin 數量

ggplot(iris, aes(Sepal.Width)) +
  geom_histogram(bins = 5)+
  labs(title = "set bins = 15")
```

## boxplot

```{r box}
ggplot(iris, aes(x=Sepal.Width)) +
  #stat_boxplot(geom = "errorbar", width =.3)+
  geom_boxplot() +
  labs(title = "boxplot of Sepal.Width",
       x = "Sepal.Width(cm)") + 
  theme_classic()+
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line.y = element_blank()) #把y軸刪掉

```

## side by side box plot

```{r box2}
ggplot(iris, aes(x=Sepal.Width, y = Species)) +
  geom_boxplot(aes(fill = Species), show.legend = F)+
  labs(title = "boxplots of Sepal.Width by Species")
```

### Using `graphic` functions

```{r base boxplot}
boxplot(Sepal.Width ~ Species, data =iris, horizontal = T)
```

## Scatter plot

```{r}
g1 <- ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width))
g1 + geom_point()
g1 + geom_point(aes(color = Species, shape = Species)) +
  labs(title = "Scatter plot of Sepal.Width, Sepal.Length by Species")
```

## Lastly, back to `graphic` plotting functions

-   One can easily draw a plot without constructing a `data.frame` object.

```{r baser}
# par(mfrow = c(2, 1))
barplot(table(iris$Species), ylab="count", main = "barplot")
hist(iris$Sepal.Width, breaks = 20, main = "historgram",xlim = c(2,4.5))
boxplot(iris$Sepal.Width, main = "boxplot", horizontal = T)
plot(iris$Sepal.Width, iris$Sepal.Length, main = "scatterplot")
```

# Exercise 

```{r}
sca <- ggplot(iris, aes(Petal.Length, Petal.Width)) +
  geom_point(aes(color = Species, shape = Species),#size =2,
             position = position_jitter(seed = 87))
sca + labs(title = "張瑋宸", tag = "A")

# ggsave("figures/hw.png", plot = sca)
```
